---
title: "MakingDataFile"
author: "Whitney Domico"
date: "2022-10-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


importing data from sources
```{r}
library(readxl)
PMC3077530_Data <- read.csv("C:/Users/Whitn/OneDrive/Desktop/BIO247/BIO247Project/BIO247ProjectArticles/PMC3077530_Data.txt", sep="", header = TRUE)
PMC2775422_Data <- read.csv("C:/Users/Whitn/OneDrive/Desktop/BIO247/BIO247Project/BIO247ProjectArticles/PMC2775422_Data.txt", sep="", header = TRUE)
PMC3912837_Data <- read_excel("C:/Users/Whitn/OneDrive/Desktop/BIO247/BIO247Project/BIO247ProjectArticles/PMC3912837_Data.xlsx",sheet = "Sheet2")
PMC2890845_Data <- read_excel("C:/Users/Whitn/OneDrive/Desktop/BIO247/BIO247Project/BIO247ProjectArticles/PMC2890845_Data.xlsx")
PMC4724864_Data <- read_excel("C:/Users/Whitn/OneDrive/Desktop/BIO247/BIO247Project/BIO247ProjectArticles/PMC4724864_Data.xlsx", skip = 1)
PMC6927206_Data <- read_excel("C:/Users/Whitn/OneDrive/Desktop/BIO247/BIO247Project/BIO247ProjectArticles/PMC6927206_Data.xlsx", skip = 1)
PMC3872086_Data <- read_excel("C:/Users/Whitn/OneDrive/Desktop/BIO247/BIO247Project/BIO247ProjectArticles/PMC3872086_Data.xlsx", skip = 5)
PMC4059435_Data <- read_excel("C:/Users/Whitn/OneDrive/Desktop/BIO247/BIO247Project/BIO247ProjectArticles/PMC4059435_Data.xlsx")
PMC3905728_Data <- read_excel("C:/Users/Whitn/OneDrive/Desktop/BIO247/BIO247Project/BIO247ProjectArticles/PMC3905728_Data.xlsx", sheet = "Sheet1")
PMC3827979_Data <- read_excel("C:/Users/Whitn/OneDrive/Desktop/BIO247/BIO247Project/BIO247ProjectArticles/PMC3827979_Data.xlsx")
```


removing letters from end of PMC3077530$SNP data and combining with other data
```{r}
temp <- substr(PMC3077530_Data$SNP,1,nchar(PMC3077530_Data$SNP)-3)

temp3 <- c()
for (each in PMC3912837_Data$SNP){
  temp2 <- unlist(strsplit(each, ""))
  if ("*" %in% temp2){
   temp3 <- c(temp3, (substr(each, 2, nchar(each))))
  } else
    temp3 <- c(temp3, each)
}
SNP <- c(temp, PMC2775422_Data$SNP, temp3, PMC2890845_Data$`Tag SNP ID`, PMC4724864_Data$SNP, PMC6927206_Data$SNP...1, PMC3872086_Data$SNP, PMC4059435_Data$SNP, PMC3905728_Data$`Strongest SNP`, PMC3827979_Data$`rs ID`)
```


splitting PMC2890845 and PMC3827979 into Chr and position
```{r}
temp4 <- c()
for (each in PMC2890845_Data$CNVR) {
  temp <- unlist(strsplit(each, ""))
  temp2 <- c(temp[4:5])
  if (":" %in% temp2 && "X" %in% temp2 == FALSE && "Y" %in% temp2 == FALSE && "x" %in% temp2 == FALSE && "y" %in% temp2 == FALSE){
    temp3 <- paste0(0, temp2[1])
    temp4 <- c(temp4, temp3)
  } else if (":" %in% temp2){
    temp4 <- c(temp4, temp2[1])
  } else
    temp5 <- paste0(temp2[1], temp2[2])
    temp4 <- c(temp4, temp5)
temp <- c()
temp2 <- c()
temp3 <- c()
temp5 <- c()
}
PMC2890845_Data$Chr <- temp4

temp4 <- c()
for (each in PMC2890845_Data$CNVR) {
  temp <- unlist(strsplit(each, ":"))
  temp2 <- c(temp[2])
  temp3 <- unlist(strsplit(temp2, "–"))
  temp4 <- c(temp4, as.numeric(temp3[1]))
  temp <- c()
  temp2 <- c()
  temp3 <- c()
}
PMC2890845_Data$Position.bp <- temp4

temp4 <- c()
for (each in PMC3827979_Data$ChrPos) {
  temp <- unlist(strsplit(each, ""))
  temp2 <- c(temp[4:5])
  if (":" %in% temp2 && "X" %in% temp2 == FALSE && "Y" %in% temp2 == FALSE && "x" %in% temp2 == FALSE && "y" %in% temp2 == FALSE){
    temp3 <- paste0(0, temp2[1])
    temp4 <- c(temp4, temp3)
  } else if (":" %in% temp2){
    temp4 <- c(temp4, temp2[1])
  } else
    temp5 <- paste0(temp2[1], temp2[2])
    temp4 <- c(temp4, temp5)
temp <- c()
temp2 <- c()
temp3 <- c()
temp5 <- c()
}
PMC3827979_Data$Chr <- temp4



```


putting positions into equal units
```{r}
temp <- PMC2775422_Data$Position.bp.
temp <- temp/1000000
temp2 <- PMC3912837_Data$Position.bp/1000000
temp3 <- PMC2890845_Data$Position.bp/1000000
temp5 <- PMC4724864_Data$BP/1000000
temp6 <- PMC6927206_Data$BP...3/1000000
temp7 <- PMC4059435_Data$`Location (bp)`/1000000
temp8 <- as.numeric(PMC3905728_Data$`Position (hg18)`)/1000000
temp9 <- PMC3827979_Data$bp/1000000
Position.Mb <- c(PMC3077530_Data$Position.Mb., temp, temp2, temp3, temp5, temp6, PMC3872086_Data$Mb, temp7, temp8, temp9)
```


removing junk characters from Chr for PMC3912837; redefining 1 as 01, etc in PMC3077530 and defining Chr
```{r}
temp4 <- c()
for (each in PMC3912837_Data$Chr) {
  temp <- unlist(strsplit(each, ""))
  temp2 <- c(temp[4:5])
  if (":" %in% temp2 && "X" %in% temp2 == FALSE && "Y" %in% temp2 == FALSE && "x" %in% temp2 == FALSE && "y" %in% temp2 == FALSE){
    temp3 <- paste0(0, temp2[1])
    temp4 <- c(temp4, temp3)
  } else if (":" %in% temp2){
    temp4 <- c(temp4, temp2[1])
  } else
    temp5 <- paste0(temp2[1], temp2[2])
    temp4 <- c(temp4, temp5)
temp <- c()
temp2 <- c()
temp3 <- c()
temp5 <- c()
}


Chr <- c()
temp6 <- c(PMC3077530_Data$Chr, PMC2775422_Data$Chr, temp4, PMC2890845_Data$Chr, PMC4724864_Data$CHR, PMC6927206_Data$CHR, PMC3872086_Data$Chr, PMC4059435_Data$Chr., PMC3905728_Data$Chromosome, PMC3827979_Data$Chr)
for (each in temp6){
  temp7 <- unlist(strsplit(each, ""))
  if (length(temp7) == 1 && "X" %in% temp7 == FALSE && "Y" %in% temp7 == FALSE && "x" %in% temp7 == FALSE && "y" %in% temp7 == FALSE) {
    temp8 <- paste0(0, each)
    Chr <- c(Chr, temp8)
 } else 
    Chr <- c(Chr, each)
temp6 <- c()
}
```


inserting PMCID columns for all data; defining PMCID
```{r}
PMC2890845_Data$PMCID <- c("PMC2890845")
PMC3912837_Data$PMCID <- c("PMC3912837")
PMC2775422_Data$PMCID <- c("PMC2775422")
PMC3077530_Data$PMCID <- c("PMC3077530")
PMC4724864_Data$PMCID <- c("PMC4724864")
PMC6927206_Data$PMCID <- c("PMC6927206")
PMC3872086_Data$PMCID <- c("PMC3872086")
PMC4059435_Data$PMCID <- c("PMC4059435")
PMC3905728_Data$PMCID <- c("PMC3905728")
PMC3827979_Data$PMCID <- c("PMC3827979")

PMCID <- c(PMC3077530_Data$PMCID, PMC2775422_Data$PMCID, PMC3912837_Data$PMCID, PMC2890845_Data$PMCID, PMC4724864_Data$PMCID, PMC6927206_Data$PMCID, PMC3872086_Data$PMCID, PMC4059435_Data$PMCID, PMC3905728_Data$PMCID, PMC3827979_Data$PMCID)
```


beginning new dataframe for all data
```{r}
InitialData <- data.frame(SNP, Chr, Position.Mb, PMCID)
```


```{r}
length(unique(InitialData$SNP))
length(InitialData$SNP)
```

making dataframe with only unique values of SNP (found setDT function on stack exchange)

```{r}
library(data.table)
require(data.table)


tempdf2 <- setDT(InitialData)[,.(Chr=paste(Chr,collapse = ",")), by=SNP]

temp2 <- c()
for (each in tempdf2$Chr){
  temp <- unlist(strsplit(each, ","))
  temp2 <- c(temp2, temp[1])
}
 
tempdf3 <- setDT(InitialData)[,.(Position.Mb=paste(Position.Mb,collapse = ",")), by=SNP]

temp3 <- c()
for (each in tempdf3$Position.Mb){
  temp <- unlist(strsplit(each, ","))
  temp3 <- c(temp3, temp[1])
}
 
tempdf4 <- setDT(InitialData)[,.(PMCID=paste(PMCID,collapse = ",")), by=SNP]

MainData <- data.frame("SNP"= tempdf2$SNP, "Chr"=temp2, "Position.Mb"=temp3, "PMCID"=tempdf4$PMCID)

```

checking to see if unique function worked

```{r}
length(unique(MainData$SNP))==length(MainData$SNP)
```
unadjusted frequency graph

```{r}
library(ggplot2)
ggplot(MainData)+geom_bar(aes(x=Chr))
```

adjusted frequency graph 

```{r}
library(readxl)
library(dplyr)
library(ggplot2)
ChrData <- read_excel("C:/Users/Whitn/OneDrive/Desktop/BIO247/BIO247Project/BIO247ProjectArticles/ChrData.xlsx")


ObsData <- as.data.frame(table(MainData$Chr))

ObsData$ChrBP <- ChrData$`Length (bp)`

ObsData$AdjObs <- 10^7*(ObsData$Freq/ObsData$ChrBP)

mean <- mean(ObsData$AdjObs)
sd <- sd(ObsData$AdjObs)

plot(x=ObsData$Var1, y=ObsData$AdjObs, xlab= "Chromosome", ylab=("Relative Frequency"), pch=19, col='black')

ObsData$AdjObs.binned <- cut(ObsData$AdjObs, c(0, 0.5*sd, 1, 1.5*sd, 2.0*sd, 3.0*sd, 3.5*sd, 4*sd, 4.5*sd, 5*sd))

ggplot(ObsData)+geom_bar(aes(x=AdjObs.binned))
```
Stat testing for adjusted frequency

```{r}
mean <- mean(ObsData$AdjObs)
sd <- sd(ObsData$AdjObs)

onesd <- 1+sd
twosd <- 1+2*sd

over68 <- c()
over95 <- c()

row <- c(1:length(ObsData$Var1))

for (each in row){
  if (ObsData[each,'AdjObs']> onesd && ObsData[each,'AdjObs'] < twosd){
      over68 <- c(over68, ObsData[each,'Var1'])
  } else if (ObsData[each,'AdjObs'] >= twosd){
      over95 <- c(over95, ObsData[each,'Var1'] )
  }
}
print(over68)
print(over95)
```

chr6 more testing

```{r}
row <- c(1:length(MainData$SNP))
chr6 <- c()
for (each in row){
  if (MainData[each,2]=="06"){
    chr6 <- c(chr6, as.numeric(MainData[each,3]))
  }
}

library(ggplot2)
testdf <- data.frame("initial"= chr6)
testdf$binnedchr6 <- cut(chr6, c(0, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140, 150, 160, 170))
ggplot(testdf)+geom_bar(aes(x=binnedchr6))


row <- c(1:length(MainData$SNP))
chr6 <- c()
SNP6 <- c()
for (each in row){
  if (MainData[each,2]=="06"){
    if (MainData[each,3]>20 && MainData[each,3]<=30){
      chr6 <- c(chr6, as.numeric(MainData[each,3]))
      SNP6 <- c(SNP6, MainData[each, 'SNP'])
    }
  }
}
data.frame("SNP"=SNP6,"Position.bp"=chr6)

library(readxl)
library(IRanges)
chr6gene <- read_excel("C:/Users/Whitn/OneDrive/Desktop/BIO247/BIO247Project/BIO247ProjectArticles/CHR6TABLE.xlsx",range = "a1:d65")
chr6gene$endposapprox <- chr6gene$`Begin position in sequence` + 12000
chr6gene$pos.Mb <- chr6gene$`Begin position in sequence`/1000000
chr6gene$end.Mb <- chr6gene$endposapprox/1000000

chr6pos <- IRanges(start=min(chr6), end=max(chr6))
chr6genes <- IRanges(start=chr6gene$pos.Mb, end=chr6gene$end.Mb)

findOverlaps(chr6pos, chr6genes)



```


chr22 more testing

```{r}
row <- c(1:length(MainData$SNP))
chr22 <- c()
for (each in row){
  if (MainData[each,2]=="22"){
    chr22 <- c(chr22, as.numeric(MainData[each,3]))
  }
}
library(ggplot2)
testdf <- data.frame("initial"= chr22)
testdf$binnedchr22 <- cut(chr22, c(0, 10, 20, 30, 40, 50))
ggplot(testdf)+geom_bar(aes(x=binnedchr22))
```
chr11 more testing

```{r}
row <- c(1:length(MainData$SNP))
chr11 <- c()
for (each in row){
  if (MainData[each,2]=="11"){
    chr11 <- c(chr11, as.numeric(MainData[each,3]))
  }
}
library(ggplot2)
testdf <- data.frame("initial"= chr11)
testdf$binnedchr11 <- cut(chr11, c(0, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140))
ggplot(testdf)+geom_bar(aes(x=binnedchr11))
```
```


