---
title: "MakingDataFile"
author: "Whitney Domico"
date: "2022-10-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

importing data from sources
```{r}
library(readxl)
PMC3077530_Data <- read.csv("C:/Users/Whitn/OneDrive/Desktop/BIO247/BIO247Project/BIO247ProjectArticles/PMC3077530_Data.txt", sep="", header = TRUE)
PMC2775422_Data <- read.csv("C:/Users/Whitn/OneDrive/Desktop/BIO247/BIO247Project/BIO247ProjectArticles/PMC2775422_Data.txt", sep="", header = TRUE)
PMC3912837_Data <- read_excel("C:/Users/Whitn/OneDrive/Desktop/BIO247/BIO247Project/BIO247ProjectArticles/PMC3912837_Data.xlsx",sheet = "Sheet2")
PMC2890845_Data <- read_excel("C:/Users/Whitn/OneDrive/Desktop/BIO247/BIO247Project/BIO247ProjectArticles/PMC2890845_Data.xlsx")
PMC3603134_Data <- read_excel("C:/Users/Whitn/OneDrive/Desktop/BIO247/BIO247Project/BIO247ProjectArticles/PMC3603134_Data.xlsx", skip = 3)
PMC4724864_Data <- read_excel("C:/Users/Whitn/OneDrive/Desktop/BIO247/BIO247Project/BIO247ProjectArticles/PMC4724864_Data.xlsx", skip = 1)
```


removing letters from end of PMC3077530$SNP data and combining with other data
```{r}
temp <- substr(PMC3077530_Data$SNP,1,nchar(PMC3077530_Data$SNP)-3)

temp3 <- c()
for (each in PMC3912837_Data$SNP){
  temp2 <- unlist(strsplit(each, ""))
  if ("*" %in% temp2){
   temp3 <- c(temp3, (substr(each, 2, nchar(each))))
  } else
    temp3 <- c(temp3, each)
}
PMC3603134_Data$SNP <- c("-")
SNP <- c(temp, PMC2775422_Data$SNP, temp3, PMC2890845_Data$`Tag SNP ID`, PMC3603134_Data$SNP, PMC4724864_Data$SNP)
```

splitting PMC2890845 and PMC3603134 CNVR into Chr and position
```{r}
temp4 <- c()
for (each in PMC2890845_Data$CNVR) {
  temp <- unlist(strsplit(each, ""))
  temp2 <- c(temp[4:5])
  if (":" %in% temp2 && "X" %in% temp2 == FALSE && "Y" %in% temp2 == FALSE && "x" %in% temp2 == FALSE && "y" %in% temp2 == FALSE){
    temp3 <- paste0(0, temp2[1])
    temp4 <- c(temp4, temp3)
  } else if (":" %in% temp2){
    temp4 <- c(temp4, temp2[1])
  } else
    temp5 <- paste0(temp2[1], temp2[2])
    temp4 <- c(temp4, temp5)
temp <- c()
temp2 <- c()
temp3 <- c()
temp5 <- c()
}
PMC2890845_Data$Chr <- temp4

temp4 <- c()
for (each in PMC2890845_Data$CNVR) {
  temp <- unlist(strsplit(each, ":"))
  temp2 <- c(temp[2])
  temp3 <- unlist(strsplit(temp2, "–"))
  temp4 <- c(temp4, as.numeric(temp3[1]))
  temp <- c()
  temp2 <- c()
  temp3 <- c()
}
PMC2890845_Data$Position.bp <- temp4

temp4 <- c()
for (each in PMC3603134_Data$...2) {
  temp <- unlist(strsplit(each, ""))
  temp2 <- c(temp[4:5])
  if (":" %in% temp2 && "X" %in% temp2 == FALSE && "Y" %in% temp2 == FALSE && "x" %in% temp2 == FALSE && "y" %in% temp2 == FALSE){
    temp3 <- paste0(0, temp2[1])
    temp4 <- c(temp4, temp3)
  } else if (":" %in% temp2){
    temp4 <- c(temp4, temp2[1])
  } else
    temp5 <- paste0(temp2[1], temp2[2])
    temp4 <- c(temp4, temp5)
temp <- c()
temp2 <- c()
temp3 <- c()
temp5 <- c()
}
PMC3603134_Data$Chr <- temp4

temp4 <- c()
for (each in PMC3603134_Data$...2) {
  temp <- unlist(strsplit(each, ":"))
  temp2 <- c(temp[2])
  temp3 <- unlist(strsplit(temp2, "-"))
  temp4 <- c(temp4, as.numeric(temp3[1]))
  temp <- c()
  temp2 <- c()
  temp3 <- c()
}
PMC3603134_Data$Position.bp <- temp4
```


putting positions into equal units
```{r}
temp <- PMC2775422_Data$Position.bp.
temp <- temp/1000000
temp2 <- PMC3912837_Data$Position.bp/1000000
temp3 <- PMC2890845_Data$Position.bp/1000000
temp4 <- PMC3603134_Data$Position.bp/1000000
temp5 <- PMC4724864_Data$BP/1000000
Position.Mb <- c(PMC3077530_Data$Position.Mb., temp, temp2, temp3, temp4, temp5)
```

removing junk characters from Chr for PMC3912837; redefining 1 as 01, etc in PMC3077530 and defining Chr

```{r}
temp4 <- c()
for (each in PMC3912837_Data$Chr) {
  temp <- unlist(strsplit(each, ""))
  temp2 <- c(temp[4:5])
  if (":" %in% temp2 && "X" %in% temp2 == FALSE && "Y" %in% temp2 == FALSE && "x" %in% temp2 == FALSE && "y" %in% temp2 == FALSE){
    temp3 <- paste0(0, temp2[1])
    temp4 <- c(temp4, temp3)
  } else if (":" %in% temp2){
    temp4 <- c(temp4, temp2[1])
  } else
    temp5 <- paste0(temp2[1], temp2[2])
    temp4 <- c(temp4, temp5)
temp <- c()
temp2 <- c()
temp3 <- c()
temp5 <- c()
}


Chr <- c()
temp6 <- c(PMC3077530_Data$Chr, PMC2775422_Data$Chr, temp4, PMC2890845_Data$Chr, PMC3603134_Data$Chr, PMC4724864_Data$CHR)
for (each in temp6){
  temp7 <- unlist(strsplit(each, ""))
  if (length(temp7) == 1 && "X" %in% temp7 == FALSE && "Y" %in% temp7 == FALSE && "x" %in% temp7 == FALSE && "y" %in% temp7 == FALSE) {
    temp8 <- paste0(0, each)
    Chr <- c(Chr, temp8)
 } else 
    Chr <- c(Chr, each)
temp6 <- c()
}
```

inserting PMCID columns for all data; defining PMCID

```{r}
PMC2890845_Data$PMCID <- c("PMC2890845")
PMC3912837_Data$PMCID <- c("PMC3912837")
PMC2775422_Data$PMCID <- c("PMC2775422")
PMC3077530_Data$PMCID <- c("PMC3077530")
PMC3603134_Data$PMCID <- c("PMC3603134")
PMC4724864_Data$PMCID <- c("PMC4724864")

PMCID <- c(PMC3077530_Data$PMCID, PMC2775422_Data$PMCID, PMC3912837_Data$PMCID, PMC2890845_Data$PMCID, PMC3603134_Data$PMCID, PMC4724864_Data$PMCID)
```


beginning new dataframe for all data
```{r}
MainData <- data.frame(SNP, Chr, Position.Mb, PMCID)
View(MainData)
```

graph for fun
```{r}
library(ggplot2)
ggplot(MainData)+geom_bar(aes(x=Chr))
```


```{r}
length(unique(MainData$SNP))
length(MainData$SNP)
```




