---
title: "mainanalysis"
author: "Whitney Domico"
date: '`r Sys.Date()`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

adding chr data to gene data -- Chr22 has NO GENES and will be removed from now on
```{r}
Chr11Genes$chr <- 11
Chr6Genes$chr <- "06"
```

combining chr 6, 11 dfs
```{r}
genedata <- data.frame("chr"=c(Chr6Genes$chr, Chr11Genes$chr))
genedata$bp <- c(Chr6Genes$potSNP6, Chr11Genes$potSNP11)
genedata$potgene <- c(Chr6Genes$potgenes6, Chr11Genes$potgenes11)
```

adding SNPs to df
```{r}
row <- c(1:length(MainData$Position.Mb))

temp <- c()
for (each in row){
  for (each2 in genedata$bp){
    if (each2 == MainData$Position.Mb[each]){
      temp <- c(temp, MainData$SNP[each])
    }
  }
}


genedata$SNP <- temp
```



adding descriptions to data

```{r}
NCBIGeneDescrData <- read_excel("C:/Users/Whitn/OneDrive/Desktop/BIO247/BIO247Project/BIO247ProjectArticles/NCBIGeneDescrData.xlsx")

genedata$officialsymbol <- NCBIGeneDescrData$officialsymbol
genedata$fullname <- NCBIGeneDescrData$fullname
genedata$description <- NCBIGeneDescrData$description
genedata$simpdesc <- NCBIGeneDescrData$simple
genedata$diseases <- NCBIGeneDescrData$diseases
```

uploading pathways data and adding to genedata
```{r}
library(readxl)
BAT2path <- read_excel("C:/Users/Whitn/OneDrive/Desktop/BIO247/BIO247Project/BIO247ProjectArticles/BAT2path.xlsx")
HLADRB1path <- HLADRB1path <- read_excel("C:/Users/Whitn/OneDrive/Desktop/BIO247/BIO247Project/BIO247ProjectArticles/HLADRB1path.xlsx", sheet = "Sheet2")
LRRC16Apath <- read_excel("C:/Users/Whitn/OneDrive/Desktop/BIO247/BIO247Project/BIO247ProjectArticles/LRRC16Apath.xlsx")
SLC17A1path <- read_excel("C:/Users/Whitn/OneDrive/Desktop/BIO247/BIO247Project/BIO247ProjectArticles/SLC17A1path.xlsx")
PPP1R10path <- read_excel("C:/Users/Whitn/OneDrive/Desktop/BIO247/BIO247Project/BIO247ProjectArticles/PPP1R10path.xlsx")
LTApath <- read_excel("C:/Users/Whitn/OneDrive/Desktop/BIO247/BIO247Project/BIO247ProjectArticles/LTApath.xlsx")
SLC44A4path <- read_excel("C:/Users/Whitn/OneDrive/Desktop/BIO247/BIO247Project/BIO247ProjectArticles/SLC44A4path.xlsx")
ZKSCAN3path <- read_excel("C:/Users/Whitn/OneDrive/Desktop/BIO247/BIO247Project/BIO247ProjectArticles/ZKSCAN3path.xlsx")
ZKSCAN12path <- read_excel("C:/Users/Whitn/OneDrive/Desktop/BIO247/BIO247Project/BIO247ProjectArticles/ZKSCAN12path.xlsx")
OR5T2path <- read_excel("C:/Users/Whitn/OneDrive/Desktop/BIO247/BIO247Project/BIO247ProjectArticles/OR5T2path.xlsx")
SMTNL1path <- read_excel("C:/Users/Whitn/OneDrive/Desktop/BIO247/BIO247Project/BIO247ProjectArticles/SMTNL1path.xlsx")
CCDC15path <- read_excel("C:/Users/Whitn/OneDrive/Desktop/BIO247/BIO247Project/BIO247ProjectArticles/CCDC15path.xlsx")
TMEM218path <- read_excel("C:/Users/Whitn/OneDrive/Desktop/BIO247/BIO247Project/BIO247ProjectArticles/TMEM218path.xlsx")
genedata$pathways <- c(BAT2path, HLADRB1path, LRRC16Apath, SLC17A1path, PPP1R10path, LTApath, SLC44A4path, ZKSCAN3path, ZKSCAN12path, OR5T2path, SMTNL1path, CCDC15path, TMEM218path)

```


cleaning pathways data
```{r}
temp <- c()
pathways <- c()
for (each in genedata$pathways){
  temp <- unlist(strsplit(each, ","))
  pathways <- c(pathways, temp)
}
pathways <- na.omit(pathways)
```

defining new df for pathways
```{r}
df <- as.data.frame(table(pathways))
```

defining new df for diseases
```{r}
diseasesdf<- (as.data.frame(table(unlist(strsplit(tolower(genedata$diseases), ",")))))
write_xlsx(diseasesdf, "C:/Users/Whitn/OneDrive/Desktop/BIO247/BIO247Project/BIO247ProjectArticles/diseasesdf.xlsx")
```

REQUIRES USER INTERACTION
open diseasesdf.xlsx; define new column "Immune"; if disease is autoimmune, put "1" in row; if not autoimmune, put "0" in row
```{r}
diseasesdfnew <- read_excel("C:/Users/Whitn/OneDrive/Desktop/BIO247/BIO247Project/BIO247ProjectArticles/diseasesdfrevised.xlsx")
```

calculating percent of autoimmune diseases in group
```{r}
immune <- c()
nonimmune <- c()
row <- 1:length(diseasesdfnew$Immune)
for (each in row){
  if (diseasesdfnew$Immune[each]==1){
    immune <- c(immune, diseasesdfnew$Var1[each])
  } else {
    nonimmune <- c(nonimmune, diseasesdfnew$Var1[each])
  }
}

length(immune)/(length(immune)+length(nonimmune))
```
importing data about Chr6 and cutting df to only include histocompatibility genes

```{r}
chr6table <- read_excel("C:/Users/Whitn/OneDrive/Desktop/BIO247/BIO247Project/BIO247ProjectArticles/CHR6TABLE.xlsx")

chr6table <- chr6table[8:20,]
chr6table$`Begin position in sequence` <- chr6table$`Begin position in sequence`/1000000
```

finding percent of chr6 SNPs that are in histocompatibility complex (assuming avg gene length of 10000 bps)
```{r}
avglength <- 10
histoSNP <- c()
nonhistoSNP <- c()
for (each in chr6){
  if (each>=chr6table$`Begin position in sequence`[1] && each<=(chr6table$`Begin position in sequence`[length(chr6table$`Begin position in sequence`)]+avglength)){
    histoSNP <- c(histoSNP, each)
  } else {
    nonhistoSNP <- c(nonhistoSNP, each)
  }
}

histoSNP <- sort(histoSNP)
nonhistoSNP <- sort(nonhistoSNP)

length(histoSNP)/(length(nonhistoSNP)+length(histoSNP))
```
adding data about which pathways are in each gene
```{r}
vect <- c()
for (each in df$pathways){
    if (each %in% BAT2path$pathways){
      vect <- c(vect, ", BAT2")
    } else
      vect <- c(vect, "no")
}
vect2 <- c()
for (each in df$pathways){
    if (each %in% HLADRB1path$pathways){
      vect2 <- c(vect2, "HLADRB1")
    } else
      vect2 <- c(vect2, "no")
}
path <- paste(vect, vect2, sep=", ")

vect2 <- c()
for (each in df$pathways){
    if (each %in% LRRC16Apath$pathways){
      vect2 <- c(vect2, "LRRC16A")
    } else
      vect2 <- c(vect2, "no")
}
path <- paste(path, vect2, sep=", ")

vect2 <- c()
for (each in df$pathways){
    if (each %in% SLC17A1path$pathways){
      vect2 <- c(vect2, "SLC17A1")
    } else
      vect2 <- c(vect2, "no")
}
path <- paste(path, vect2, sep=", ")

vect2 <- c()
for (each in df$pathways){
    if (each %in% PPP1R10path$pathways){
      vect2 <- c(vect2, "PPP1R10")
    } else
      vect2 <- c(vect2, "no")
}
path <- paste(path, vect2, sep=", ")

vect2 <- c()
for (each in df$pathways){
    if (each %in% LTApath$pathways){
      vect2 <- c(vect2, "LTA")
    } else
      vect2 <- c(vect2, "no")
}
path <- paste(path, vect2, sep=", ")

vect2 <- c()
for (each in df$pathways){
    if (each %in% SLC44A4path$pathways){
      vect2 <- c(vect2, "SLC44A4")
    } else
      vect2 <- c(vect2, "no")
}
path <- paste(path, vect2, sep=", ")

vect2 <- c()
for (each in df$pathways){
    if (each %in% ZKSCAN3path$pathways){
      vect2 <- c(vect2, "ZKSCAN3")
    } else
      vect2 <- c(vect2, "no")
}
path <- paste(path, vect2, sep=", ")

vect2 <- c()
for (each in df$pathways){
    if (each %in% ZKSCAN12path$pathways){
      vect2 <- c(vect2, "ZKSCAN12")
    } else
      vect2 <- c(vect2, "no")
}
path <- paste(path, vect2, sep=", ")

vect2 <- c()
for (each in df$pathways){
    if (each %in% OR5T2path$pathways){
      vect2 <- c(vect2, "OR5T2")
    } else
      vect2 <- c(vect2, "no")
}
path <- paste(path, vect2, sep=", ")

vect2 <- c()
for (each in df$pathways){
    if (each %in% SMTNL1path$pathways){
      vect2 <- c(vect2, "SMTNL1")
    } else
      vect2 <- c(vect2, "no")
}
path <- paste(path, vect2, sep=", ")

vect2 <- c()
for (each in df$pathways){
    if (each %in% CCDC15path$pathways){
      vect2 <- c(vect2, "CCDC15")
    } else
      vect2 <- c(vect2, "no")
}
path <- paste(path, vect2, sep=", ")

vect2 <- c()
for (each in df$pathways){
    if (each %in% TMEM218path$pathways){
      vect2 <- c(vect2, "TMEM218")
    } else
      vect2 <- c(vect2, "no")
}
path <- paste(path, vect2, sep=", ")

temp <- c()

temp <- gsub("no", "", path)

temp2 <- gsub(", ,",",",temp)
temp3 <- gsub(", ,",",",temp2)
temp4 <- gsub(", ,",",",temp3)
temp5 <- gsub(", ,",",",temp4)

temp6 <- sub(".{2}", "", temp5)
temp7 <- sub(".{2}$", "", temp6)

df$genes <- temp7
```

adding percent column

```{r}
df$percent <- df$Freq/length(genedata$potgene)*100
```


adding plot for amounts of pathways that are present in certain percents of genes (pathways in only 1 gene have been ommitted)
```{r}
percents <- c()

for (each in df$Freq){
  if (each >1){
    percents <- c(percents, each)
  }
}
percents <- percents/length(genedata$potgene)*100
percents <- as.data.frame(percents)

ggplot(percents)+ geom_bar(aes(x=percents))
```
finding common words in pathway titles
```{r}
rem <- c()
commonwords <- c()
temp <- c()
temp <- unlist(strsplit(pathways, " "))
temp <- gsub("\\(", "", temp)
temp <- gsub("\\)", "", temp)
temp <- gsub("\\[", "", temp)
temp <- gsub("\\]", "", temp)
temp <- gsub("/", "", temp)
temp <- gsub("\\\\", "", temp)
temp <- gsub("\\-", "", temp)
temp <- na.omit(temp)

for (each in temp){
  temp2 <- unlist(strsplit(each, ""))
  if (length(temp2)<4 && each != "RNA" && each != "HPA" && each != "MHC" && each != "CD4"){
    rem <- c(rem, each)
  } else {
    commonwords <- c(commonwords, each)
  }
}
finaldata <- as.data.frame(table(tolower((commonwords))))

finaldata <- finaldata[order(-finaldata$Freq),]
```

saving an excel sheet with word list and frequencies
```{r}
library(writexl)
finaldata$relfreq <- finaldata$Freq/max(finaldata$Freq)
write_xlsx(finaldata, "C:/Users/Whitn/OneDrive/Desktop/BIO247/BIO247Project/BIO247ProjectArticles/finaldata.xlsx")
```

finding significant words and redefining data to only include these
```{r}
mean <- mean(finaldata$Freq)
sd <- sd(finaldata$Freq)

onesd <- mean+sd
twosd <- mean+2*sd

over68 <- c()
over95 <- c()

row <- c(1:length(finaldata$Var1))

for (each in row){
  if (finaldata[each,'Freq']> onesd && finaldata[each,'Freq'] < twosd){
      over68 <- c(over68, finaldata[each,'Var1'])
  } else if (finaldata[each,'Freq'] >= twosd){
      over95 <- c(over95, finaldata[each,'Var1'] )
  }
}
relevant <- length(over68)+length(over95)


finaldata <- finaldata[1:relevant,]
```

REQUIRES USER INTERACTION
open finaldata.xlsx; name new column "kept"; insert "1" for words to continue in analysis and "0" for words to omit from analysis; save file as finaldatarevised.xlsx
```{r}
finaldatanew <- read_excel("C:/Users/Whitn/OneDrive/Desktop/BIO247/BIO247Project/BIO247ProjectArticles/finaldatarevised.xlsx")
```

removing "0" words
```{r}
row <- length(finaldatanew$kept):1

for (each in row){
  if (finaldatanew$kept[each] == 0){
    finaldatanew <- finaldatanew[-c(each),]
  }
}

finaldata$numforgraph <- 1:length(finaldata$Var1)


temp <- c()
row <- 1:length(finaldata$Var1)
for (each in finaldatanew$Var1){
  for (each2 in row){
    if (each == finaldata$Var1[each2]){
      temp <- c(temp, finaldata$numforgraph[each2])
    }
  }
}
finaldatanew$numforgraph <- temp

```

plots for (1) counts of all significant words and (2) counts of all chosen words

```{r}
library(ggplot2)

ggplot(finaldata, color= station, fill='black', aes(x=Var1, y=Freq))+
  geom_col(position='dodge')+
  ylab("Frequency")+
  xlab("Word")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.25, hjust=1))

ggplot(finaldatanew, color= station, fill='black', aes(x=Var1, y=Freq))+
  geom_col(position='dodge')+
  ylab("Frequency")+
  xlab("Word")
```






