---
title: "EulerProblems"
author: "Whitney Domico"
date: "2022-10-08"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

"By considering the terms in the Fibonacci sequence whose values do not exceed four million, find the sum of the even-valued terms." (answer = 4613732 )

```{r}  
RS <- c(0)
prev <- c(1)
fib <- c(2)

while (fib < 4000000){
  if (fib %% 2 == 0){
    RS <- RS + fib
    temp <- fib
    fib <- prev + fib
    prev <- temp
  } else {
    temp <- fib
    fib <- fib + prev
    prev <- temp
  }
}

print(RS)
```  


"Find the largest palindrome made from the product of two 3-digit numbers." (answer = 906609)

```{r}
vec1 <- (999:100)
pal <- c(0)

for (each in vec1){
  for (each2 in vec1){
 temp <- each*each2
 temp2 <- as.numeric(strsplit(as.character(temp),"")[[1]])
 temp3 <-(temp2[1:length(temp2)] == temp2[length(temp2):1])
 if (sum(temp3 == TRUE)==length(temp3)){
    if (temp[1]>pal[1]){
      pal <- temp
    }
 }
  }
}
    
print(pal) 
```


"What is the smallest positive number that is evenly divisible by all of the numbers from 1 to 20?" WORKS BUT INEFFICIENT (answer = 232792560)

```{r}
vec4 <- (20000000:232792600)
vec5 <- (1:20)
x <- 0
stop <- FALSE
answer2 <- 111111111110

while (stop == FALSE){
  for (each in vec4){
    for (each2 in vec5){
      x <- each %% each2 + x
      }
    if (x == 0){
      if(each < answer2){
      answer2 <- each
      }
      stop <- TRUE
    }
    x <- 0
  }
}

print(answer2)
```
```{r}
vec4 <- (100:10000)
vec5 <- (1:10)
x <- 0
stop <- FALSE

while (stop == FALSE)
```

"Find the difference between the sum of the squares of the first one hundred natural numbers and the square of the sum." (answer = 25164150)

```{r}
vec1 <- 1:100
temp <- c()

temp1 <- sum(vec1^2)
temp2 <- (sum(vec1))^2

answer<-temp2-temp1

print(answer)

```

"What is the 10 001st prime number?" DOES NOT WORK (answer = 104743)

```{r}
counter <- 0
vec1 <- c(2:150000)
x <- 0
is.wholenumber <- function(x, tol = .Machine$double.eps^0.5) {
    abs(x - round(x)) < tol
}


for (each in vec1){
  if (each %% 2 != 0 || each == 2){
    x <- x + 1
    if (each %% 3 != 0 || each == 3){     
      x <- x + 1
      if (each %% 4 != 0 || each == 4){
        x <- x + 1
        if (each %% 5 != 0 || each == 5){
          x <- x + 1
          if (each %% 6 != 0 || each == 6){
            x <- x + 1
            if (each %% 7 != 0 || each == 7){
              x <- x + 1
              if (each %% 8 != 0 || each == 8){
                x <- x + 1
                if (each %% 9 != 0 || each == 9){
                  x <- x + 1
                  if (each %% 10 != 0 || each == 10){
                    x <- x + 1
                    if (each %% 11 != 0 || each == 11){
                      x <- x + 1
                      if (each %% 12 != 0 || each == 12){
                        x <- x + 1
                      if (each %% 13 != 0 || each == 13){
                        x <- x + 1
                      if (x == 12 && is.wholenumber(sqrt(each)) == FALSE){
                        counter <- counter + 1
                        if (counter == 10001){
                          prime <- each
                          print(prime)
                        }
                        }
                        }
                      }
                    }
                  }
                }
              }
            }
          }
        }
      }
    }
  }
  x <- 0
}

          
        
```

"What is the 10 001st prime number?" WORKS BUT WHY? (answer = 104743)

```{r}
totalPrimes <- 0;
currentPrime <- 0;
currentNumber <- 0;


while (totalPrimes <= 10001) {
  foundADividend <- FALSE;
  
  currentNumber <- currentNumber + 1;
  
  
  for (x in 2:(as.integer(currentNumber/2)+1)) {
    if (currentNumber %% x == 0 && currentNumber != x)
    {
      foundADividend <- TRUE;
      break;
    }
  }
  
  if (foundADividend == FALSE) {
    totalPrimes <- totalPrimes + 1;
    currentPrime <- currentNumber;
    
  }  
  
}

```

"There exists exactly one Pythagorean triplet for which a + b + c = 1000.
Find the product abc." (answer = 31875000)

```{r}
vecab <- c(1:700)
vecc <- c(1)
answer <- c()

for (each in vecab){
  for (each2 in vecab){
    vecc <- sqrt(each^2 + each2^2)
    if (vecc[1]+each+each2 == 1000){
      product <- each*each2*vecc
      print(product)
    }
  }
}


```

"Find the sum of all the multiples of 3 or 5 below 1000." (answer = 233168)

```{r}
vec1 <- 3*(1:333)
vec2 <- 5*(1:199)
vec3 <- 15*(1:66)

answer <- sum(vec1) + sum(vec2) - sum(vec3)

print(answer)

```

"Find the sum of the digits in the number 100!" (answer = 648)

```{r}

vec1 <- format(prod(100:1), scientific = FALSE)

vec1 <- as.numeric(strsplit(as.character(vec1),"")[[1]])

vec2 <- sum(vec1)

print(vec2)

```

