---
title: "R Day4"
output: html_document
date: "2022-09-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Bioinformatics -- biology, statistics, computer science

robustness -- have they considered a third variable?

reproducibility -- can this be copied?

get files for this rmd

```{r}
Test_V_output <- read.csv("C:\\Users\\Whitn\\OneDrive\\Desktop\\BIO247\\BIO247Lab\\BIO247Lab3\\Test_V_output.csv")
TestData <- read.csv("C:\\Users\\Whitn\\OneDrive\\Desktop\\BIO247\\BIO247Lab\\BIO247Lab2\\Dataset_S1.txt")
```

## Exercise 1: Rows and Columns

The goal for this exercise is to save the Total Scores column into a vector and then find the 'Total Score' and molecules involved in the interaction for the user's "favorite row," which for these purposes will be defined as row 17.

The code below creates a vector of the total scores.

```{r include=TRUE}
tot_score <- c(Test_V_output$Total.Score)
```

To find the molecules involved in row 17, one must find both the 'Target.Name' and 'Source.Name'.

The code below calls the 'Target.Name' involved in row 17.

```{r include=TRUE}
Test_V_output[17,'Target.Name']
```

The code below calls the 'Source.Name' involved in row 17.

```{r include=TRUE}
Test_V_output[17,'Source.Name']
```

Together, these comprise the molecules involved. Therefore, the answer is tsc2 and ppargamma.

The final part of this exercise asks for the Total Score of row 17, which uses the code below.

```{r include=TRUE}
Test_V_output[17,'Total.Score']
```

Therefore, the Total Score for row 17 is 240.

## Exercise 2: Mean and Stddev

The goal of this exercise is to find the mean and standard deviation of the Total Scores in Test_V\_output. The second part is to find how many unique papers were used in this set.

To find the mean of the Total Scores, the following code is used. Recall that a previous vector titled 'tot_score' was created

```{r include=TRUE}
mean(tot_score)
```

Therefore, the mean is approximately 63.8.

To find the standard deviation of the Total Scores, the following code is used.

```{r include=TRUE}
sd(tot_score)
```

Therefore, the standard deviation is approximately 55.3.

The second part of this exercise is to find the amount of unique papers that were used in this set. These would be found under the 'Paper ID' column and would be coded as follows.

```{r include=TRUE}
length(unique(Test_V_output$Paper.ID))
```

Therefore, the number of unique papers is 157. Of course, this counts groups of papers as separate items.If we were to find the true unique number of papers, we would first have to separate the items in the vector into singular papers. The first step to this is to remove all "junk characters," which are "[", "]", " ", "'".

```{r include=TRUE}
Test_V_output$Paper.ID <- gsub("\\'","", Test_V_output$Paper.ID)
Test_V_output$Paper.ID <- gsub("\\[", "", Test_V_output$Paper.ID)
Test_V_output$Paper.ID <- gsub("\\]", "", Test_V_output$Paper.ID)
Test_V_output$Paper.ID <- gsub(" ","", Test_V_output$Paper.ID)
```

Then the individual papers need to be separated. This can be achieved by splitting them where there are commas. This will return a list (not a vector), so we also must use the 'unlist' function to create the new vector, SplitPaperIDs.

```{r include=TRUE}
SplitPaperIDs <- unlist(strsplit(Test_V_output$Paper.ID, ","))
```

Now that the vector is created, the length can be found using the 'unique' function.

```{r include=TRUE}
length(unique(SplitPaperIDs))
```

Therefore, the true number of unique papers is 102.

## Exercise 3: Vectorization

The goal of this exercise is to recreate the Total Score column by calculating it using the evidence, kind, match, and epistemic scores. It is given that Total Score = (Kind+(Evidence(Match))(Epistemic).

The vector names for kind, evidence, match, and epistemic scores are as follows, respectively:

-   *Test_V\_output\$Kind.Score*

-   *Test_V\_output\$Evidence.Score*

-   *Test_V\_output\$Match.Score*

-   *Test_V\_output\$Epistemic.Value*

The new column will be named "New.Total.Score", and is coded for as follows:

```{r include=TRUE}
Test_V_output$New.Total.Score <- (Test_V_output$Kind.Score+(Test_V_output$Evidence.Score*Test_V_output$Match.Score))*Test_V_output$Epistemic.Value
```

## Exercise 4: Subsets

The goal of this exercise is to remove the columns labeled "Location", "Cell.Type", and "Organism", and also to view the dataframe in a way that only includes the first 100 rows.

To remove the aforementioned columns, the following code is used:

```{r include=TRUE}
Test_V_output = subset(Test_V_output, select = -c(Location, Cell.Type, Organism))
```

To view the dataframe as only the first 100 rows, the following code is used (it is represented as a store function to avoid the output of 100 lines):

```{r include=TRUE}
store <- Test_V_output[1:100,]
```

## Indexing
list
```{r}
word<-"tacocat"
substr(word,1,1)
substr(word,1,4)
substr(word,1,7)
```

```{r}
word <- c("t","a","c","o","c","a","t")
word == word [length(word):1]
```

## If/Else Statements

```{r echo=FALSE}
#shirt<-"red"
#shirt<-"yellow"
#shirt<-"green"
#shirt<-"blue"
shirt<-"white"
```
^^ this is to choose the value of 'shirt'
```{r echo=FALSE}
if (shirt=="red") {
  print("Eat apple") 
} else if (shirt=="yellow"){
  print("Eat banana")
} else if (shirt=="green"){
  print("Eat spinach")
} else if (shirt=="blue"){
  print ("Eat blueberries")
} else {
  print ("Eat yogurt")
} 
```

## While Functions

```{r echo=FALSE}
num <- 6
while (num<20){
  num <- num+1
}
```

Example 

```{r}
stop <- sample (1:1000,1)
num <- 1
while (num<stop){
  print("...")
  num <- num+1
}
print("Pop!")
```

## For Functions

```{r echo=FALSE}
vect1 <- c(1,2,3)
for (each in vect1){
  print(each)
}
```

## Graphs

to plot a dataframe column density

```{r include=TRUE}
ggplot(Test_V_output)+ geom_density(aes(x=Total.Score), fill="black")
```

to plot a dataframe column density with specific restraints -- %in% = only; >, <, !=

```{r include=TRUE}
ggplot(Test_V_output[Test_V_output$Total.Score > 140,])+ geom_density(aes(x=Total.Score), fill="black")
ggplot(TestData[TestData$total.SNPs >70,])+geom_point(aes(x=total.SNPs, y=total.Bases))
```

to plot a bar chart

```{r}
ggplot(TestData) + geom_bar(aes(x=Divergence.binned2))
```

## Binning

new column name <- cut(column to be binned, number of bins needed)
```{r include=TRUE}
TestData$Divergence.binned <- cut(TestData$Divergence, 5)
```
OR
new column name <- cut(column to be binned, c(x, y, z))

```{r include=TRUE}
TestData$Divergence.binned2 <- cut(TestData$Divergence, c(0, 0.02, 0.04, 0.06, 0.08, 0.10, 0.12, 1.6))
ggplot(TestData)+geom_bar(aes(Divergence.binned2))
```

## Counting Values

find values within perimeters of a certain vector; sum != length
```{r}
sum(Test_V_output$Total.Score < 240) + sum(Test_V_output$Total.Score >=240) == length(Test_V_output$Total.Score)
sum(Test_V_output$Total.Score)
length(Test_V_output$Total.Score)
```

## Coding Diagrams

For unique paper ID problem

want -- how many unique papers are there in the PaperID colum in Test_V_output?

Note: This is a vector of lists.

1. Remove "junk" characters
  gsub("[", "]", "//', " ", df$PaperIDs)
2. Split up lists of IDs
  SplitPaperIDs <- unlist(strsplit(df$PaperIDs, ","))
3. Revectorize to vector of individual IDs
  vect1 = c()
  for (every in SplitPaperIDs){
    append to vect1
  }
4. Count unique paper IDs
  length(unique(vect1))
  
```{r}  
RS <- c(0)
prev <- c(1)
fib <- c(2)

while (fib < 4000000){
  if (fib %% 2 == 0){
    RS <- RS + fib
    temp <- fib
    fib <- prev + fib
    prev <- temp
  } else {
    temp <- fib
    fib <- fib + prev
    prev <- temp
  }
}


```  

```{r}
vec1 <- (999:100)
pal <- c(0)

for (each in vec1){
  for (each2 in vec1){
 temp <- each*each2
 temp2 <- as.numeric(strsplit(as.character(temp),"")[[1]])
 temp3 <-(temp2[1:length(temp2)] == temp2[length(temp2):1])
 if (sum(temp3 == TRUE)==length(temp3)){
    if (temp[1]>pal[1]){
      pal <- temp
      print(pal)
    }
 }
  }
}
    
  
```
```{r}
vec4 <- (20000000:232792600)
vec5 <- (1:20)
x <- 0
stop <- FALSE
answer2 <- 111111111110

while (stop == FALSE){
  for (each in vec4){
    for (each2 in vec5){
      x <- each %% each2 + x
      }
    if (x == 0){
      if(each < answer2){
      answer2 <- each
      }
      stop <- TRUE
    }
    x <- 0
  }
}

print(answer2)
```
```{r}
vec4 <- (100:10000)
vec5 <- (1:10)
x <- 0
stop <- FALSE

while (stop == FALSE)
```


```{R}
y<-2520
for (each in 1:10){
  print(y %% each)
}
```
